### 2024-05-01 Paper/paper.tex
Updates from Overleaf

\captionsetup{justification=raggedright,singlelinecheck=false}

@@

	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance Programs}

@@

    First, I test empirically whether nonprofits respond to program incentives differently from for-profits. While nonprofit hospital objectives are largely unknown, for-profits have revealed their objective in their ownership classification. That is, $\alpha=1$. To learn something about $\alpha$ for nonprofit firms with different leadership compositions, I compare their program response to that of for-profits. I estimate the following regression equation:

    \begin{equation}
    \label{eq:forprofit}
    y_{ht} = \beta (for\_profit_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the penalties took place, and $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively. For each outcome, I estimate three models where the comparison group differs by hospital type: all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive.

    An important distinction here is whether hospital time is a time-invariant characterization. For nonprofits and for-profits, switching is relatively rare, and I limit to those who never switch. However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period can change. Under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), these changes can be advantageous in understanding hospital objectives. There are two ways that having a clinically trained executive might affect outcomes: first, a hospital that hires a clinically trained executive is signalling their underlying $\alpha$. Second, hiring a clinically trained executive who carries out day-to-day operations in a specific way actually changes the underlying $\alpha$ of the hospital. Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.

    Similarly, I test empirically whether having a clinically trained executive affects hospital response to program incentives. Thus, I limit the sample to only hospitals for which I have leadership team information, and estimate:

    \begin{equation}
    \label{eq:clinical}
    y_{ht} = \beta (clinical_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}.
    \end{equation}

@@

    Before presenting the main findings, I investigate whether executive team changes occur endogenously due to the programs. If hospitals change their propensity to hire a clinically trained executive because they expect to be penalized or want to gain extra payments, then there is selection in the independent variable. First, I analyze whether changes in clinically trained executives are concentrated around the enactment of both HRRP and HVBP, 2011-2012. Second, I analyze whether hospitals who end up getting penalized through HRRP are more likely to change clinically trained executives on their team than non-penalized hospitals, and similarly for hospitals who end up getting extra payments for high quality of care through HVBP. That is, are hospitals that expect to be more affected by the programs more likely to hire or fire clinically trained executives? I consider changes along having any clinically trained executive as well as changes in the number of clinically trained executives. I estimate the following regression equations:

@@

    \begin{equation}
    change_{ht} = \beta(ever\_penalized_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
    \end{equation}

    The estimates from this analysis are presented in Table \ref{tab:change_analysis}. Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executives are actually becoming more stable after the programs. Further, hospitals who end up being penalized are not more likely to change their propensity to hire any clinically trained executive. Results are similar when considering changes in the number of clinically trained executives. Thus, I conclude that changes in executive teams are not endogenously occurring with the change in incentives due to the programs.

@@

     First, I present results for the differential program response of for-profit firms compared to the different sub-samples of nonprofit firms in Table \ref{tab:forprofit_synth}. Looking first at the left panel, for-profits change their readmission rates after the program similarly to all nonprofits taken together, as well as nonprofits without clinically trained executives. However, for-profits have a .3ppt larger decrease in readmissions than nonprofits with clinically trained executives. When considering the right panel, there is no statistical difference in change in mortality rates between different hospital types.

@@

     Next, I directly estimate the difference in response by nonprofits with and without clinically trained executives. These results are presented in Table \ref{tab:MD_noMD_synth}. Nonprofits without clinically trained executives have a larger decrease in readmissions than those with clinically trained executive by .3ppts, while there is no difference in change in mortality rates.

     These findings are consistent with the theoretical prediction presented in Section \ref{sec:model} if having a clinically trained executive means the hospital places more weight on societal benefit than for-profits and other nonprofits do.

@@

@@

    As a secondary analysis, I present estimates of differential response of potential mechanisms, changing input into another societal benefit measure: uncompensated care, and the complexity of patients seen in the hospital: case mix index. First, I investigate whether for-profits change these differentially than types of nonprofits in Table \ref{tab:forprofit_uncompCMI_synth}.

>>>>

\captionsetup{justification=center,singlelinecheck=false}

@@

	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance Incentives}

@@

    First, I test empirically whether different sub-samples of nonprofits respond to program incentives differently from for-profits. While nonprofit hospital objectives are largely unknown, for-profits have revealed their objective in their ownership classification. That is, $\alpha=1$. To learn something about $\alpha$ for nonprofit firms with different executive team compositions, I compare their program response to that of for-profits. I estimate the following regression equation:

    \begin{equation}
    \label{eq:forprofit}
    y_{ht} = \beta (for\_profit_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the penalties took place, and $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively. For each outcome, I estimate three models where the comparison group differs by hospital type: all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive. I also directly estimate the difference in response between hospital with and without a clinically trained executive:

    \begin{equation}
    \label{eq:clinical}
    y_{ht} = \beta (no\_MD\_exec_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

    An important distinction here is whether the sub-categorization of nonprofit hospitals by executive team is time-invariant. For nonprofits and for-profits, switching overall ownership type is rare, and I limit to those who never switch. However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change. Therefore, I first present effects of having a clinically trained executive among stable executive teams. That is, among hospitals that don't change their propensity to hire a clinically trained executive over time.

    However, under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), using the changes can actually be advantageous in better understanding underlying hospital motives. There are two reasons we might see having a clinically trained executive affecting outcomes: first, a clinically trained executive is a signal of the underlying weight placed on profit, $\alpha$. Second, hiring a clinically trained executive who carries out day-to-day operations in a specific way actually changes the underlying $\alpha$. Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place. I use the same empirical specification with these two variables as the independent variation in executive teams.

@@

    Before presenting the main findings, I investigate whether executive team changes occur endogenously due to the programs. If hospitals change their propensity to hire a clinically trained executive because they expect to be penalized or want to gain extra payments, then there is selection in the independent variable. First, I analyze whether changes in clinically trained executives are concentrated around the enactment of both HRRP and HVBP, 2011-2012. Second, I analyze whether hospitals who end up getting penalized through HRRP are more likely to change clinically trained executives on their team than non-penalized hospitals, and similarly for hospitals who end up getting extra payments for high quality of care through HVBP. That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives? I consider changes along having any clinically trained executive as well as changes in the number of clinically trained executives. I estimate the following regression equations:

@@

    \begin{equation}
    change_{ht} = \beta(program\_exposed_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
    \end{equation}

    The variable $program\_exposed_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP. The estimates from this analysis are presented in Table \ref{tab:change_analysis}. Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executives are actually becoming more stable over time along the dimension of hiring clinically trained executives. Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive. Results are similar when considering changes in the number of clinically trained executives. Thus, I conclude that changes in executive teams are likely not endogenously occurring with the incentive changes due to the programs.

@@

     First, I consider the differential program response of for-profit firms compared to the different sub-samples of nonprofit firms. These results are presented in Table \ref{tab:forprofit_synth}. Models in the left panel estimate the effect of being for-profit after the programs were enacted, compared to nonprofit sub-samples, on readmission rates. There is no statistically meaningful difference between the response of for-profits and nonprofits as a whole, or nonprofits without a clinically trained executive. However, for-profits lower their readmissions by .3ppts compared to nonprofits without clinically trained executives. Models in the right panel estimate the effect of being for-profit after the programs on mortality rates. Here, there is no statistical difference in mortality rate response between different hospital types.

@@

     Next, I directly estimate the difference in response by nonprofits with and without clinically trained executives. These results are presented in Table \ref{tab:MD_noMD_synth}, where the left has models with outcome readmission rates and the right has models with outcome mortality rates. Nonprofits without clinically trained executives have a larger decrease in readmissions than those with clinically trained executive by .3ppts, while there is no difference in change in mortality rates. Tables \ref{tab:forprofit_synth} and \ref{tab:MD_noMD_synth} taken together show that, in terms of readmission rates, nonprofits without clinically trained executives behave more similarly to for-profits than they do other nonprofits. Under the assumption that having a clinically trained executive on the leadership team indicates a higher preference for social utility, these findings are consistent with the theoretical prediction presented in Section \ref{sec:model}.

@@

    However, it is not clear whether these results are due to clinically trained executives being a signal of underlying hospital objectives, or that the leaders themselves make a difference. Therefore, I present a decomposition of the results using two models in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either whether the hospital ever has a clinically trained executive, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted. The left panel on readmission rates shows that the difference is response is fully driven by having a clinically trained executive in 2012 compared to having a clinically trained trained executive at another point in time. Again, there is nothing going on with mortality.

@@

    As a secondary analysis, I present estimates of differential response of potential mechanisms. The first is changing input into another societal benefit measure: uncompensated care, and the second is complexity of patients: case mix index. First, I investigate whether for-profits change these differentially than types of nonprofits in Table \ref{tab:forprofit_uncompCMI_synth}.

### 2024-05-01 Tables/MD_noMD_readmort_synth.tex
Updates from Overleaf

\caption{Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Weighted Avg. Readmission Rate} & \multicolumn{1}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2)\\
\midrule
NP w/ MD Executive x Post Penalty & -0.28$^{*}$ & -0.19\\
 & (0.12) & (0.11)\\
\textbf{Comparison Group:} &  & \\
NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
\\
Observations & 2,335 & 2,325\\
\bottomrule
\end{tabular}
\end{table}

>>>>

\caption{\label{tab:MD_noMD_synth}Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Weighted Avg. Readmission Rate} & \multicolumn{1}{c}{Weighted Avg. Mortality Rate} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
 & (1) & (2)\\
\midrule
NP w/out MD Exec x Post Penalty & -0.28$^{*}$ & -0.19\\
 & (0.12) & (0.11)\\
\textbf{Comparison Group:} &  & \\
NP w/ MD Executive & $\checkmark$ & $\checkmark$\\
Observations & 2335 & 2325\\
\bottomrule
\end{tabular}
\end{table}
